### 2020-08-26 paper/theory/model_executive_bargain.tex
update theory section
incorporate new edits to the theory section

\begin{document}

@@

In this section, I adapt the canonical vote-buying model proposed by Groseclose and Snyder (1996) and Banks (2000) to model how executive control over the city council determines patronage. The theoretical framework centers on two main actors, a government (\emph{governo}) and opposition (\emph{oposição}) who compete for votes in the city council.\footnote{I use the terms mayor and government interchangeably for the rest of this section.} The mayor seeks to enact her preferred policy, but can only do so by passing a simple majority vote in the council. The opposition, on the other hand, wants to block her efforts and maintain the status quo. In order to build a winning coalition, the mayor must spend a key political resource: patronage appointments into the municipal bureaucracy.

Ultimately, two forces shape the mayor's strategy: the amount of resources controlled by the opposition and the underlying preferences of the legislators. The key parameter of interest is the underlying preferences -- i.e. support for the mayor -- of city councilors. In some municipalities a mayor has widespread political support ("base política"), while in others, the opposition ("oposição") garners more seats. In the actual empirical estimation, we approximate these preferences by looking into the councilor's partisan alignments in the mayoral election. For the second determinant, the opposition's resource, it is clear that an opposition with more resources can better compete with the mayor, forcing the latter to spend more to deter them.

I show that, in equilibrium, the cost of passing the mayor's preferred policy decreases monotonically as the legislative support for the mayor increases. Substantively, the model predicts that municipalities in which the mayor has more (less) councilors favorable to her, we should observe less (more) patronage. In the remainder of this section, I outline the setting of the model and derive key comparative statics that guide the empirical estimation of this paper. I conclude with a discussion of the model and how it applies to the Brazilian municipal context.

@@

The government $G$ and opposition $O$ compete over legislative votes to enact their preferred policies. There are two possible outcomes: a policy $x$ favored by the government, and the status quo, denote as $y$, preferred by the opposition. In order to implement her policy the mayor must pass a simple majority vote in the city council, comprised of an odd number $N$ of legislators. The total amount of political resources available is $W_G$ and $W_O$ to win over votes, which for the mayor includes patronage appointments into the public sector.

Each city councilor is fully characterized by a publicly observed policy preference $v_i$ for all $i \in N$, where $v_i > 0$ entails that the mayor's proposal $x$ is preferred. Let $\mathbf{v} = (v_1, ..., v_n)$ describe the preference profile for the city council. In this setting, $v_i$ measures the degree to which an individual city councilor supports the mayor, with higher values of $v$ denoting stronger support for the mayor and vice versa. Payoff gets realized when city councilor $i$ votes, independent of the outcome of the voting procedure.

@@

\begin{enumerate}
    \item Government $G$ offers a bribe schedule $a \in (a_1, ..., a_n) \in \real$.
    \item Opposition $O$ observes the bribe schedule $m$ and makes a counter-offer $b \in (b_1, ..., b_n) \in \real$.
    \item City councilors cast their votes at the end of bribing period.
    \item Nature gathers votes and payoffs are realized.
\end{enumerate}

@@

>>>>

\bibliographystyle{unsrtnat}
\begin{document}

@@

In this section, I use a canonical vote-buying model in legislatures \citep{groseclose_1996_buying, banks_2000_buying} to model mayor's strategy to use patronage to buy legislative votes. For any given electoral term, a mayor wants to enact her preferred policy, but to do so needs to pass it through a simple majority in the city council. To guarantee her victory, the mayor offers appointment positions to voters, in this case, city councilors, who in exchange vote in favor of the bill.

The amount of patronage the mayor offers responds to the composition of the city council. When city councilors are predominantly in favor of the mayor's policy, we should expect less patronage, and vice versa. The intuition is straightforward: if the city councilors are already likely to support the bill, there is no additional need to offer them additional side payments. Note that there is no claim about the presence or complete absence of patronage \citep{stokes_brokers_2013, grindle_jobs_2012}, but rather its degrees: how many appointments must the mayor give out in order to guarantee her victory at the voting of the policy.

An important feature of the model is the presence of an opposition party that competes with the mayor for legislative votes. In this setting, passing a policy proposal requires outbidding the competitor, which generates interesting results regarding the optimal size of the coalition and conditions under which we should observe minimal, super and universal coalitions. In particular, the amount of resources that the opposition controls can force the mayor's hands and increase the amount of patronage necessary to pass the bill. We pursue these empirical implicaitons of the model further in our results section.

For the remainder of this section, I outline the setting of the model and derive key comparative statics that guide the empirical estimation of this paper. I conclude with a discussion of the model and how it applies to the Brazilian municipal context.

@@

The mayor's government $G$ and opposition $O$ compete over legislative votes to enact their preferred policies. There are two possible outcomes: a policy $x$ favored by the government, and the status quo, denote as $y$, preferred by the opposition. In order to implement her policy the mayor must pass a simple majority vote in the city council, comprised of an odd number $N$ of legislators. The total amount of political resources available is $W_G$ and $W_O$, which for the mayor includes patronage appointments into the public sector.

Each city councilor is fully characterized by a publicly observed policy preference $v_i$ for all $i \in N$, where $v_i > 0$ entails that the mayor's proposal $x$ is preferred. Let $\mathbf{v} = (v_1, ..., v_n)$ describe the preference profile for the city council. In this setting, $v_i$ measures the degree to which an individual city councilor supports the mayor, with higher values of $v$ denoting stronger support for the mayor and vice versa. Payoff are realized when city councilor $i$ votes, independent of the outcome of the voting procedure. This sincere voting preference forecloses the possibility of general equilibria in which $i$'s voting behavior affect $j$.

@@

\begin{enumerate}
    \item Government $G$ offers a bribe schedule $a \in (a_1, ..., a_n) \in \real$.
    \item Opposition $O$ observes the bribe schedule $m$ and makes a counter-offer $b \in (b_1, ..., b_n) \in \real$.
    \item City councilors cast their votes at the end of bribing period.
    \item Nature sums legislative votes, legislative outcome is decided and payoffs are realized.
\end{enumerate}

@@

\bibliography{../dissertation}
